---
title: "Leaflet Map"
author: "Yadder Aceituno"
date: "November 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

For this assignment I had to create a webpage where I can show how to plot a map with Leaflet (library). I was able to use any dataset so I decided to find something about earthquakes occurrences around the world. The place that I go when I need data was Kaggle, and I found an interesting competition about the relation between earhquakes and solar system objects. The dataset used in that competition is used here to complete my assingment.


## Libraries

Lets import the libraries

```{r library, warning=FALSE}
library("leaflet")
library("dplyr")
```


## Data Source

You can find the research on this link: https://www.kaggle.com/aradzhabov/earthquakes-solar-system-objects/download

Lets make a first look for the data.

```{r first_look_1}

dfDataSource <- read.csv("SolarSystemAndEarthquakes.csv")


# Lets se how many variables the dataframe has.
ncol(dfDataSource)

```

As we can see, the dataframe has many variables To make the plot we will use only variables called as "earthquake".


```{r first_look_2}
columnsEarhquakes <- colnames(dfDataSource)[grepl("earthquake", colnames(dfDataSource))]
dfEarthquakes <- dfDataSource[, columnsEarhquakes]

# Summary for dataframe
summary(dfEarthquakes)

```

## Preparing Data

First, we need to create a classification variable according magnitude.

```{r preparing_1}

dfEarthquakes <- dfEarthquakes %>% 
                mutate(earthquake.classification = case_when(
                  (earthquake.mag >= 8) ~ "Great",
                  (earthquake.mag >= 7) ~ "Major",
                  (earthquake.mag >= 6) ~ "Strong",
                  (earthquake.mag >= 5) ~ "Moderate",
                  (earthquake.mag >= 4) ~ "Light",
                  TRUE ~ "Minor"
                ))

table(dfEarthquakes$earthquake.classification)
```

Now, lets cast the earthquake.time variable as date and then lets create a label variable.

```{r preparing_2}
# Casting to date
dfEarthquakes$earthquake.time <- as.Date(dfEarthquakes$earthquake.time)

# Creating a label variable
dfEarthquakes$label <- paste(as.character(dfEarthquakes$earthquake.time), " | Magnitude:", dfEarthquakes$earthquake.mag, sep = "")

```

## Plotting


Now we already have our dataframe to make the leaflet plot. I decided to make a plot where we can see the earthquake ocurrences.

```{r plot, warningFALSE}

dfQuakes <- data.frame(
  lat = dfEarthquakes$earthquake.latitude,
  lng = dfEarthquakes$earthquake.longitude,
  cls = dfEarthquakes$earthquake.classification,
  lbl = dfEarthquakes$label
)

dfGreatQuakes <- dfQuakes[dfQuakes$cls == "Great",]
dfMajorQuakes <- dfQuakes[dfQuakes$cls == "Major",]
dfStrongQuakes <- dfQuakes[dfQuakes$cls == "Strong",]

greatIcon <- awesomeIcons(
   icon = 'exclamation-triangle',
   iconColor = 'white',
   library = 'fa',
   markerColor = 'red'
 )

majorIcon <- greatIcon
majorIcon$markerColor <- 'orange'

strongIcon <- greatIcon
strongIcon$markerColor <- 'blue'


lfletMap <- leaflet() %>% addTiles() %>% 
  
  addAwesomeMarkers(data = dfGreatQuakes, 
                    lat = ~lat, lng = ~lng,
                    group = "Great", 
                    icon = greatIcon, 
                    label = ~lbl,
                    clusterOptions = markerClusterOptions()) %>%
  
  addAwesomeMarkers(data = dfMajorQuakes, 
                    lat = ~lat, lng = ~lng,
                    group = "Major", 
                    icon = majorIcon,
                    label = ~lbl,
                    clusterOptions = markerClusterOptions()) %>%
  
  addAwesomeMarkers(data = dfStrongQuakes, 
                    lat = ~lat, lng = ~lng,
                    group = "Strong", 
                    icon = strongIcon,
                    label = ~lbl,
                    clusterOptions = markerClusterOptions()) %>%
  
  addLayersControl(overlayGroups = c("Great", "Major", "Strong"),
                   options = layersControlOptions(collapsed = FALSE));

lfletMap

```
